### 2019-09-08 ic-journal/coarsest.tex
Merged all revise suggestions from Sangiorgi

The coarsest congruence
contained in (weak) bisimilarity, namely \emph{bisimilarity
  congruence}, is
the \emph{composition closure} (\texttt{CC}) of (weak) bisimilarity:
\begin{alltt}
   \HOLConst{WEAK_CONGR} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{CC} \HOLConst{WEAK_EQUIV}\hfill{[WEAK_CONGR]}
   \HOLConst{CC} \HOLFreeVar{R} \HOLSymConst{\HOLTokenDefEquality{}} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{g} \HOLBoundVar{h}. \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{c}. \HOLConst{CONTEXT} \HOLBoundVar{c} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} \ensuremath{(}\HOLBoundVar{c} \HOLBoundVar{g}\ensuremath{)} \ensuremath{(}\HOLBoundVar{c} \HOLBoundVar{h}\ensuremath{)}\ensuremath{)}\hfill{[CC_def]}
\end{alltt}
We do not need to put $R\ g\ h$ in the antecedent of
  \texttt{CC\_def}, as this is anyhow obtained from the trivial context $(\lambda x.\,x)$.
For any \hl{binary} relation $R$ 
on CCS processes, the composition closure of $R$ is always at least as
fine as $R$
(here $\subseteq_r$ stands for \emph{relational subset})
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R}. \HOLConst{CC} \HOLBoundVar{R} \HOLSymConst{\HOLTokenRSubset{}} \HOLBoundVar{R}\hfill{[CC_is_finer]}
\end{alltt}
Furthermore, we prove that any (pre)congruence contained in $R$ (which
itself may not be) is contained in the composition closure of $R$
(hence the closure is the coarsest one):
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R} \HOLBoundVar{R\sp{\prime}}. \HOLConst{congruence} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenRSubset{}} \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenRSubset{}} \HOLConst{CC} \HOLBoundVar{R}\hfill{[CC_is_coarsest]}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R} \HOLBoundVar{R\sp{\prime}}. \HOLConst{precongruence} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenRSubset{}} \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenRSubset{}} \HOLConst{CC} \HOLBoundVar{R}\hfill{[PCC_is_coarsest]}
\end{alltt}

@@

The formal proofs  of the above two results  follow Milner's proof
\cite{Mil89}. The assumption is Milner's result 
requires  $\mathrm{fn}(p) \cup
\mathrm{fn}(q) \neq \mathscr{L}$ (where $\mathrm{fn}$ stands for \emph{free
  names}), and is therefore strictly stronger than the assumption in
our result above, where we only look at immediate weak
derivatives $p$ and $q$, requiring that there is at least one action
that never appears as a label; moreover such an action need not be
the same for $p$ and for $q$
%  If only $p$ (or $q$) has free actions while the other uses all available
% labels, the assumption $\mathrm{fn}(p) \cup
% \mathrm{fn}(q) \neq \mathscr{L}$ (here $\mathrm{fn}$ stands for \emph{free
%   names}) does not hold, and the proof cannot be completed. Our
% assumption is \emph{weaker}
% in the sense that $p$ and $q$ need not have the
% \emph{same} free actions 
(and where an action $a$ and its complementary  $\outC a$ are
considered \emph{different} actions).

@@

\subsection{Arbitrary many non-bisimilar processes}

The assumption in the previous \texttt{PROP3_COMMON} requires the
existence of a special CCS process, which is not weakly bisimilar to
any sub-process emanating from the two root processes by weak
transitions. In the worst case, there may be infinite such
sub-processes (even on finitely branching processes).
We can however consider  the equivalence classes of CCS processes 
modulo weak bisimilarity.
If there are infinitely many of such classes, 
then it will be 
possible to choose one  that is distinct from all the (finitely many) states in the
graphs of the two given processes.  
This can be done following Klop's contruction \cite{van2005characterisation}.
 We call  ``Klop processes''  the processes in this construction:
\begin{definition}[Klop processes]
For each ordinal $\lambda$, and an arbitrary chosen non-$\tau$ action $a$,
define a CCS process $k_\lambda$ as follows:
\begin{itemize}
\item $k_0 = 0$,
\item $k_{\lambda+1} = k_\lambda + a.k_\lambda$ and
\item for $\lambda$ a limit ordinal, $k_\lambda = \sum_{\mu < \lambda}
  k_\mu$, meaning that $k_\lambda$ is constructed from all graphs
  $k_\mu$ for $\mu < \lambda$ by identifying their root.
\end{itemize}
\end{definition}
When processes are finite-state, that is,
the number of  states in which a process may evolve by performing
transitions is finite, 
we can use  the following subset of Klop processes, 
defined as a recursive function (on natural numbers) in HOL4:
\begin{definition}{(Klop processes as recursive function on natural numbers)}
\begin{alltt}
\HOLConst{KLOP} \HOLFreeVar{a} \HOLNumLit{0} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{\ensuremath{\mathbf{0}}}
\HOLConst{KLOP} \HOLFreeVar{a} \ensuremath{(}\HOLConst{SUC} \HOLFreeVar{n}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n} \HOLSymConst{\ensuremath{+}} \HOLConst{label} \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\hfill{[KLOP_def]}
\end{alltt}
\end{definition}

By induction on the definition of Klop processes,  and using the SOS inference
rules ($\mathrm{Sum}_1$) and ($\mathrm{Sum}_2$), we can  prove
the following properties of Klop functions:
\begin{proposition}{(Properties of Klop functions and processes)}
\begin{enumerate}
\item (All Klop processes are stable)
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{STABLE} \ensuremath{(}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\ensuremath{)}\hfill[KLOP_PROP0]
\end{alltt}
\item (Any transition from a Klop process leads to a smaller Klop
  process, and conversely)
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n} \HOLTokenTransBegin\HOLConst{label} \HOLFreeVar{a}\HOLTokenTransEnd \HOLFreeVar{E} \HOLSymConst{\HOLTokenEquiv{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{m}. \HOLBoundVar{m} \HOLSymConst{\HOLTokenLt{}} \HOLFreeVar{n} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{KLOP} \HOLFreeVar{a} \HOLBoundVar{m}\hfill{[KLOP_PROP1]}
\end{alltt}
\item (The weak version of the previous property)
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n} \HOLTokenWeakTransBegin\HOLConst{label} \HOLFreeVar{a}\HOLTokenWeakTransEnd \HOLFreeVar{E} \HOLSymConst{\HOLTokenEquiv{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{m}. \HOLBoundVar{m} \HOLSymConst{\HOLTokenLt{}} \HOLFreeVar{n} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{KLOP} \HOLFreeVar{a} \HOLBoundVar{m}\hfill{[KLOP_PROP1']}
\end{alltt}
\item (All Klop processes are distinct according to strong bisimilarity)
\begin{alltt}
\HOLTokenTurnstile{} \HOLFreeVar{m} \HOLSymConst{\HOLTokenLt{}} \HOLFreeVar{n} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenNeg{}}\ensuremath{(}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{m} \HOLSymConst{\HOLTokenStrongEQ} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\ensuremath{)}\hfill{[KLOP_PROP2]}
\end{alltt}
\item (All Klop processes are distinct according to weak bisimilarity)
\begin{alltt}
\HOLTokenTurnstile{} \HOLFreeVar{m} \HOLSymConst{\HOLTokenLt{}} \HOLFreeVar{n} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenNeg{}}\ensuremath{(}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{m} \HOLSymConst{\HOLTokenWeakEQ} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\ensuremath{)}\hfill{[KLOP_PROP2']}
\end{alltt}
\item (Klop functions are one-one)
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ONE_ONE} \ensuremath{(}\HOLConst{KLOP} \HOLFreeVar{a}\ensuremath{)}\hfill{[KLOP_ONE_ONE]}
\end{alltt}
\end{enumerate}
\end{proposition}
Having
 a recursive function  on all natural numbers,
  we can map these into a set containing  all Klop processes. 
We can therefore 
choose a number that is mapped onto a Klop process
that is distinct (i.e., not bisimilar)  from 
any  of the derivatives of two  given 
two finite-state CCS processes $p$ and $q$
(by definition, the number of such derivatives is finite).
This is done by 
appealing to a basic set-theory result.

>>>>

The coarsest congruence
contained in (weak) bisimilarity, namely \emph{bisimilarity
  congruence}, is
the \emph{composition closure} (\texttt{CC}) of (weak) bisimilarity:
\begin{alltt}
   \HOLConst{WEAK_CONGR} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{CC} \HOLConst{WEAK_EQUIV}\hfill{[WEAK_CONGR]}
   \HOLConst{CC} \HOLFreeVar{R} \HOLSymConst{\HOLTokenDefEquality{}} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{g} \HOLBoundVar{h}. \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{c}. \HOLConst{CONTEXT} \HOLBoundVar{c} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} \ensuremath{(}\HOLBoundVar{c} \HOLBoundVar{g}\ensuremath{)} \ensuremath{(}\HOLBoundVar{c} \HOLBoundVar{h}\ensuremath{)}\ensuremath{)}\hfill{[CC_def]}
\end{alltt}
We do not need to put $R\ g\ h$ in the antecedent of
  \texttt{CC\_def}, as this is anyhow obtained from the trivial context $(\lambda x.\,x)$.
\hl{The next result shows that, for any binary} relation $R$ 
on CCS processes, the composition closure of $R$ is always at least as
fine as $R$\hl{:} ($\subseteq_r$ stands for \emph{relational subset}.)
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R}. \HOLConst{CC} \HOLBoundVar{R} \HOLSymConst{\HOLTokenRSubset{}} \HOLBoundVar{R}\hfill{[CC_is_finer]}
\end{alltt}
Furthermore, we prove that any (pre)congruence contained in $R$
\hl{(which needs not to be itself a (pre)congruence)}
is contained in the composition closure of $R$
(hence the closure is the coarsest one):
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R} \HOLBoundVar{R\sp{\prime}}. \HOLConst{congruence} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenRSubset{}} \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenRSubset{}} \HOLConst{CC} \HOLBoundVar{R}\hfill{[CC_is_coarsest]}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{R} \HOLBoundVar{R\sp{\prime}}. \HOLConst{precongruence} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenRSubset{}} \HOLBoundVar{R} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{R\sp{\prime}} \HOLSymConst{\HOLTokenRSubset{}} \HOLConst{CC} \HOLBoundVar{R}\hfill{[PCC_is_coarsest]}
\end{alltt}

@@

The formal proofs  of the above two results follow Milner's proof~\cite{Mil89}.
The assumption \hl{in} Milner's result 
requires $\mathrm{fn}(p) \cup
\mathrm{fn}(q) \neq \mathscr{L}$ (where $\mathrm{fn}$ stands for \emph{free
  names}), and is therefore strictly stronger than the assumption in
our result above. \hl{Indeed, in out result we} only look at immediate weak
derivatives for $p$ and $q$, \hl{and only requires that there is an input
or output action}
that never appears as a label. \hl{Furthermore, such an action need
not to be the same for $p$ and for $q$,
as an action $a$ and its complementary $\outC a$ are
considered as \emph{different} actions.}

@@

\subsection{Arbitrarily many non-bisimilar processes}
\label{ss:arbitrarily}

The assumption in the previous \texttt{PROP3_COMMON} requires the
existence of a special CCS process, which is not weakly bisimilar to
any sub-process emanating from the two root processes by weak
transitions.
\hl{There could be infinitely many such sub-processes, even on finitely
branching processes.}
We can, however, consider the equivalence classes of CCS processes
modulo weak bisimilarity.
If there are infinitely many of such classes, 
then it will be 
possible to choose one  that is distinct from all the (finitely many) states in the
graphs of the two given processes.  
This can be done following Klop's contruction \cite{van2005characterisation}.
 We call  ``Klop processes''  the processes in this construction:
\begin{definition}[Klop processes]
For each ordinal $\lambda$, and an arbitrary chosen non-$\tau$ action $a$,
define a CCS process $k_\lambda$ as follows:
\begin{itemize}
\item $k_0 = 0$,
\item $k_{\lambda+1} = k_\lambda + a.k_\lambda$ and
\item for $\lambda$ a limit ordinal, $k_\lambda = \sum_{\mu < \lambda}
  k_\mu$, meaning that $k_\lambda$ is constructed from all graphs
  $k_\mu$ for $\mu < \lambda$ by identifying their root.
\end{itemize}
\end{definition}
When processes are finite-state, that is,
the number of  states in which a process may evolve by performing
transitions is finite, 
we can use  the following subset of Klop processes, 
defined as a recursive function (on natural numbers) in HOL4:
\begin{definition}{(Klop processes as recursive function on natural numbers)}
\begin{alltt}
\HOLConst{KLOP} \HOLFreeVar{a} \HOLNumLit{0} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{\ensuremath{\mathbf{0}}}
\HOLConst{KLOP} \HOLFreeVar{a} \ensuremath{(}\HOLConst{SUC} \HOLFreeVar{n}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n} \HOLSymConst{\ensuremath{+}} \HOLConst{label} \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\hfill{[KLOP_def]}
\end{alltt}
\end{definition}

\hl{Following the inductive structure of the definition of Klop processes,}
  and using the SOS rules
  ($\mathrm{Sum}_1$) and ($\mathrm{Sum}_2$), we proved
the following properties of Klop functions:
\begin{proposition}{(Properties of Klop functions and processes)}
\begin{enumerate}
\item (All Klop processes are stable)
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{STABLE} \ensuremath{(}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\ensuremath{)}\hfill[KLOP_PROP0]
\end{alltt}
\item (Any transition from a Klop process leads to a smaller Klop
  process, and conversely)
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n} \HOLTokenTransBegin\HOLConst{label} \HOLFreeVar{a}\HOLTokenTransEnd \HOLFreeVar{E} \HOLSymConst{\HOLTokenEquiv{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{m}. \HOLBoundVar{m} \HOLSymConst{\HOLTokenLt{}} \HOLFreeVar{n} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{KLOP} \HOLFreeVar{a} \HOLBoundVar{m}\hfill{[KLOP_PROP1]}
\end{alltt}
\item (The weak version of the previous property)
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n} \HOLTokenWeakTransBegin\HOLConst{label} \HOLFreeVar{a}\HOLTokenWeakTransEnd \HOLFreeVar{E} \HOLSymConst{\HOLTokenEquiv{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{m}. \HOLBoundVar{m} \HOLSymConst{\HOLTokenLt{}} \HOLFreeVar{n} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{KLOP} \HOLFreeVar{a} \HOLBoundVar{m}\hfill{[KLOP_PROP1']}
\end{alltt}
\item (All Klop processes are distinct according to strong bisimilarity)
\begin{alltt}
\HOLTokenTurnstile{} \HOLFreeVar{m} \HOLSymConst{\HOLTokenLt{}} \HOLFreeVar{n} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenNeg{}}\ensuremath{(}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{m} \HOLSymConst{\HOLTokenStrongEQ} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\ensuremath{)}\hfill{[KLOP_PROP2]}
\end{alltt}
\item (All Klop processes are distinct according to weak bisimilarity)
\begin{alltt}
\HOLTokenTurnstile{} \HOLFreeVar{m} \HOLSymConst{\HOLTokenLt{}} \HOLFreeVar{n} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenNeg{}}\ensuremath{(}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{m} \HOLSymConst{\HOLTokenWeakEQ} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\ensuremath{)}\hfill{[KLOP_PROP2']}
\end{alltt}
\item (Klop functions are one-one)
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ONE_ONE} \ensuremath{(}\HOLConst{KLOP} \HOLFreeVar{a}\ensuremath{)}\hfill{[KLOP_ONE_ONE]}
\end{alltt}
\end{enumerate}
\end{proposition}
Having
 a recursive function  on all natural numbers,
  we can map these into a set containing  all Klop processes. 
We can therefore 
choose a number that is mapped onto a Klop process
that is distinct (i.e., not bisimilar)  from 
any  of the derivatives of two  given 
two finite-state CCS processes $p$ and $q$
(by definition, the number of such derivatives is finite).
This is done by 
appealing to a basic set-theory result.

### 2019-09-08 ic-journal/contraction.tex
Merged all revise suggestions from Sangiorgi

\begin{enumerate}
\item   $P \arr\mu P'$ implies there is $Q'$ such that $Q \arcap \mu
  Q'$ and $P' \RR Q'$;
\item $Q \arr\mu Q'$   implies there is $P'$ such that $P \Arcap \mu
 P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\emph{Bisimilarity contraction}, written as $P \mcontrBIS Q$ ($P$
\emph{contracts to} $Q$), if $P\ \R\ Q$ for some contraction $\R$.
\end{definition}
% We say that $P$  {\emph{bisimilarly  contracts} $Q$, written
% $Q \mcontrBIS Q$.
% We write $\mexpaBIS$ for the inverse of $\mcontrBIS$.

In the first clause $Q$ is required to match $P$'s challenge
transition with at most one transition.
This makes sure that $Q$ is capable of mimicking $P$'s
work at least as efficiently as $P$. 
In contrast, the second clause of Definition~\ref{d:BisCon}, on the
challenges from $Q$, entirely ignores efficiency: it is the same
clause of  weak bisimulation~---the final derivatives are even required
to be related  by $\wb$, rather than by $\R$.

Bisimilarity  contraction is coarser than 
 the \emph{expansion relation} $\expa$~\cite{arun1992efficiency,sangiorgi2015equations}.
This is a
preorder widely used in proof techniques for bisimilarity and that 
intuitively refines bisimilarity by 
 formalising the idea of `efficiency' between processes.
Clause (1) is the same in the two
preorders. But in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
 moreover with
contraction the final derivatives are simply required to be bisimilar.
An expansion 
$P \expa Q$
tells us  that $Q$ is always at least as efficient as $P$, whereas  the
 contraction $P \mcontrBIS Q$  just says that $Q$ has the possibility of
being at least as efficient as $P$.

@@

\begin{definition}%[expansion]
\label{d:expa}
A process relation ${\R}$
  is an \emph{expansion} if, whenever
we have $P\RR Q$, for all $\mu$
 \begin{enumerate}
 \item   $P \arr\mu P'$ implies that there is $Q'$ with $Q \arcap \mu
   Q'$
  and $P' \RR Q'$;
 \item
     $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu
  P'$ and $P'
 \RR Q'$.
 \end{enumerate}
  $P$  {\em expands} $Q$, written as
 $P  \expa Q$,
 if $P \RR Q$ for some expansion $\R$.
 \end{definition}

@@

% The substitutivity proofs for for expansion and bisimilarity carry
% over to contraction,

Like the case of (weak) bisimilarity, the expansion and contraction
preorders are preserved by all CCS operators but the summation.

>>>>

\begin{enumerate}
\item   $P \arr\mu P'$ implies there is $Q'$ such that $Q \arcap \mu
  Q'$ and $P' \RR Q'$;
\item $Q \arr\mu Q'$   implies there is $P'$ such that $P \Arcap \mu
 P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\hl{Two processes $P$ and $Q$ are in the \emph{bisimilarity
    contraction}}, written as $P \mcontrBIS Q$ ($P$ \emph{contracts to} $Q$),
if $P\ \R\ Q$ for some contraction $\R$.
\hl{Sometimes we write $\mexpaBIS$ for the inverse of $\mcontrBIS$.}
\end{definition}

In the first clause $Q$ is required to match $P$'s challenge
transition with at most one transition.
This makes sure that $Q$ is capable of mimicking $P$'s
work at least as efficiently as $P$. 
In contrast, the second clause of Definition~\ref{d:BisCon}, on the
challenges from $Q$, entirely ignores efficiency: it is the same
clause of  weak bisimulation~---the final derivatives are \hl{indeed} required
to be related  by $\wb$, rather than by $\R$.

Bisimilarity  contraction is coarser than 
the \emph{expansion relation} $\expa$~\cite{arun1992efficiency,sangiorgi2015equations}.
This is a preorder \hl{that is} widely used in proof techniques for bisimilarity and
intuitively refines bisimilarity by 
formalising the idea of `efficiency' between processes.
Clause (1) is the same in the two
preorders. But in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
 moreover with
contraction the final derivatives are simply required to be bisimilar.
An expansion 
$P \expa Q$
tells us  that $Q$ is always at least as efficient as $P$, whereas  the
 contraction $P \mcontrBIS Q$  just says that $Q$ has the possibility of
being at least as efficient as $P$.

@@

\begin{definition}%[expansion]
\label{d:expa}
A process relation ${\R}$
  is an \emph{expansion} if, whenever
we have $P\RR Q$, for all $\mu$
 \begin{enumerate}
 \item   $P \arr\mu P'$ implies that there is $Q'$ with $Q \arcap \mu  Q'$
  and $P' \RR Q'$;
 \item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu P'$ and $P' \RR Q'$.
 \end{enumerate}
\hl{Two processes $P$ and $Q$ are in the \emph{bisimilarity
    expansion}}, written as $P \expa Q$ ($P$ \emph{expands} $Q$),
 if $P \RR Q$ for some expansion $\R$.
 \end{definition}

@@

\hl{In the same way as bisimilarity, both the expansion and the
contraction preorders are preserved by all CCS operators except the
summation. The proofs are similar, see e.g.}~\cite{sangiorgi2017equations}.

### 2019-09-08 ic-journal/equa.tex
text
Clean up the rest DDS macros; disable line numbers
Merged all revise suggestions from Sangiorgi

\begin{definition}
Assume that, for each $i$ of 
 a countable indexing set $I$, we have variables $X_i$, and expressions
$E_i$ possibly containing  such variables. 
Then 
$\{  X_i = E_i\}_{i\in I}$
is 
  a \emph{system of equations}. (There is one equation for each variable $X_i$.)
\end{definition}

@@

For instance, the solution of the equation 
$ X = a. X$ 
is  the process
$R \DSdefi \recu A {\, (a. A)}$, and for any other solution $P$ we have $P \wb R$.
In contrast, the equation 
 $X = a|  X$ has solutions that may be quite different, for instance,
 $K$ and $K | b$, for $K \DSdefi \recu K {\, (a. K)}$. (Actually any process capable of
continuously performing $a$ actions (while behaves arbitrarily on
other actions) is a solution for $X = a|  X$.)

@@

\item \emph{(strongly) guarded} if, in each $E_i$, each occurrence of
  an \behav\  variable is underneath a \emph{visible} prefix;

\item \emph{sequential} if, in each $E_i$, each of its subexpressions with occurrence of
  an \behav\  variable, apart from the variable itself, is in forms of
  prefixes or sums.
\end{itemize}
\end{definition}

% In other words,  
% if the system is sequential, then 
%  for
% every expression $E_i$, any subexpression of $E_i$ in which $X_j $ 
% appears, apart from $X_j$ itself,  is a sum (of prefixed terms). 
% For instance, 
% \begin{itemize}
% \item $X = \tau. X + \mu . \nil$ is sequential but not 
%  guarded, because the guarding prefix for the variable
% is not visible.

% \item $X =  \ell . X | P$ is  guarded but not sequential.

% \item $X =  \ell . X + \tau. \res a (a .\outC b | a.\nil)$, as well as 
% $X = \tau . (a. X + \tau . b .X + \tau  )$
% are both 
%  guarded and sequential.
% \end{itemize}

\begin{theorem}[unique solution of equations, \cite{Mil89}]
\label{t:Mil89}
A system of guarded and sequential equations (without direct sums)
$\{  X_i = E_i\}_{i\in I}$ has a unique solution for $\wb$.
\end{theorem}

% The proof exploits an invariance property on immediate transitions for
% guarded and sequential expressions, and then extracts a bisimulation
% (up-to bisimilarity) out
% of the solutions of the system.  
To see the  need of the sequentiality  condition, consider
 the equation (from \cite{Mil89}) $X = \res a (a. X | \outC a)$
where $X$ is guarded but not sequential. Any process that does not use $a$ is a solution.

>>>>

\begin{definition} % Def 3.1
Assume that, for each $i$ of 
 a countable indexing set $I$, we have variables $X_i$, and expressions
$E_i$ possibly containing  such variables. 
Then 
$\{  X_i = E_i\}_{i\in I}$
is 
  a \emph{system of equations}. (There is one equation for each variable $X_i$.)
\end{definition}

@@

For instance, the solution of the equation 
$ X = a. X$ 
is  the process
$R \DSdefi \recu A {\, (a. A)}$, and for any other solution $P$ we have $P \wb R$.
In contrast, the equation 
 $X = a|  X$ has solutions that may be quite different, for instance,
 $K$ and $K | b$, for $K \DSdefi \recu K {\, (a. K)}$. (Actually any process capable of
continuously performing \hl{$a$--actions} is a solution \hl{of} $X = a  |  X$.)

@@

\item \emph{(strongly) guarded} if, in each $E_i$, each occurrence of
  an equation variable is underneath a \emph{visible} prefix;

\item \emph{sequential} if, in each $E_i$, each \hl{occurrence of an
    equation variable only appears underneath prefixes and sums.}
\end{itemize}
\end{definition}

\hl{In other words, if the system is sequential, then for
every expression $E_i$, any subexpression of $E_i$ in which $X_j $
appears, apart from $X_j$ itself, is a sum (of prefixed terms).
For instance,}
\begin{itemize}
\item $X = \tau. X + \mu . \nil$ is sequential but not 
 guarded, because the guarding prefix for the variable
is not visible.

\item $X =  \ell . X | P$ is guarded but not sequential.

\item $X =  \ell . X + \tau. \res a (a .\outC b | a.\nil)$, as well as
$X = \tau . (a. X + \tau . b .X + \tau  )$ are both guarded and sequential.
\end{itemize}

\begin{theorem}[unique solution of equations, \cite{Mil89}]
\label{t:Mil89}
A system of guarded and sequential equations (\hl{with guarded sums only})
$\{  X_i = E_i\}_{i\in I}$ has an unique solution for $\wb$.
\end{theorem}

\hl{The proof exploits an invariance property on immediate transitions for
guarded and sequential expressions, and then extracts a bisimulation
(up to bisimilarity) out
of the solutions of the system.  }
To see the need of the sequentiality  condition, consider
 the equation (from \cite{Mil89}) $X = \res a (a. X | \outC a)$
where $X$ is guarded but not sequential. Any process that does not use
$a$ is a solution\hl{, e.g. $\nil$ and $b.\nil$.}

%% next file: contraction.tex
